Fast Radio Burst Cosmology

### 2014-04-15 writeup/dm_angular_ps.tex
Filled in some details of estimator, started Sensitivty section.

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\{D_i\}$ with known angular positions but with unknown
redshifts depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\{z_i\})$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\{z_i\}) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $\{D^{\bar{n_e}}_i\}$ are
uncorrelated:
\begin{equation}
\langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
    - \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle
    = \delta_{ij} \times \textrm{(something to do with $p(z_i)$)}
\end{equation}

@@

The covariance matrix is
\begin{align}
    C_{ij} =& \langle D_i D_j \rangle - \langle D_i \rangle \langle D_j \rangle
           = \textrm{a bunch of things we know (special case for $i=j$)}.
\end{align}

Now that we have the covariance matrix, the estimator is Equation~11.94 from
Dodelson, although we have to modify it slightly to exclude the
auto-correlations.  The error (assuming Gaussianity) is Equation~11.100.

@@

>>>>

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\{D_i\}$ with known angular positions but with unknown
redshifts depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\{z_i\})$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\{z_i\}) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $\{D^{\bar{n_e}}_i\}$ are
uncorrelated:
\begin{align}
\langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
    - \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle
    =& \delta_{ij} \int_0^\infty \ud z p(z) \int_0^{\eta(z)} \ud \eta'
        (1 + z'(\eta')) \textrm{(not quite right, subtract off the mean.)}
        \equiv& \sigma_{\bar{n}_e}^2
\end{align}

@@

The covariance matrix is
\begin{align}
    C_{ij} \equiv& \langle D_i D_j \rangle - \langle D_i \rangle \langle D_j \rangle
           =& \textrm{a bunch of things we know (special case for $i=j$)}.
           =& w(\hat{n}_i\cdot\hat{n}_j) \quad (i \neq j)
    =& \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \textrm(the $w$ term) \quad (i = j)
\end{align}

Now that we have the covariance matrix, we construct a quadratic estimator for
a chosen set of parameters $\{\lambda_\alpha\}$. The estimator is Equation~11.94 from
Dodelson, although we have to modify it slightly to exclude the
auto-correlations.  The error (assuming Gaussianity) is Equation~11.100.

The dispersion measure angular correlation function 
$w(\hat{n}_i\cdot\hat{n}_j)$ depends on the cosmology
and the redshift distribution function $p(z)$, for which we may parameterize
our ignorance. The
cross-correlations between the observed dispersion measures are thus useful
probes of cosmology.  The auto-correlations on the other hand are contaminated
by the highly uncertain $\sigma^2_{\rm src}$ as well as the $C^{zz'}_\ell$ on
very small scales and zero redshift separation, which is difficult to model
theoretically.  As such they do not carry useful information.

To explicitly ignore the information in the auto-correlations we define
\begin{equation}
    \tilde{C}_{ij,\alpha} \equiv
\begin{cases}
    C_{ij,\alpha} & \text{if } i \neq j,\\
0 & \text{if } i = j
\end{cases}.
\end{equation}
By replacing $C_{ij,\alpha}$ with this quantity in the above estimator, we
obtain an estimator which explicitly excludes information in the auto-products
and only uses the diagonal of the covariance in the noise weights. Our
ignorance of $\sigma^2_{\rm src}$ then only affects the optimality of the
estimator, although in practise
$\langle D_i D_i \rangle - \langle D_i \rangle \langle D_i \rangle$ would
simply be measured directly from the data to obtain accurate noise weights.

@@

\section{Sensitivity}

In this section we calculate the sensitivity of a hypothetical survey observing
X FRBs. Because $C_\ell$ contains the all the information we use bins
of $C_\ell$ as our parameters. We use the error bar on these parameters given
in Equation~\ref{e:error}.  We note that it is highly unlikely that an observed
set $\{D_i\}$ is even remotely Gaussian, and as such our error-bars are likely
underestimated and out estimator sub-optimal at the order unity level. 
However since given that the statistics of FRBs are currently essentially
unconstrained, we continue with this assumption.

We assume parameters X, Y, Z etc.
